### 2013-02-14 doc/icml2013/abstract.tex
updates

\begin{abstract} 
  Spectral algorithms for latent variable models have seen considerable
  recent interest for being efficient consistent estimators of model
  parameters. These algorithms make few if any assumptions about the
  generative process of the data, while providing a polynomial sample
  and computational complexity. We present a new spectral algorithm for
  a discriminative model, a mixture of linear regressions and show that
  it can recover the regression coefficients with similar polynomial
  guarantees. We evaluate the algorithm on linearly and non-linearly
  generated data, as well as on a motion tracking task and compare it's
  characteristics with an E-M algorithm.
\end{abstract}

>>>>

\begin{abstract}

Discriminative latent-variable models are typically learned using
EM or gradient-based optimization, which suffer from local optima.
In this paper, we develop a new computationally efficient
and provably consistent estimator for the mixture of linear regressions,
a simple instance of discriminative latent-variable models.
Our approach relies on a low-rank linear regression to recover
a symmetric tensor, which can be factorized into the parameters
using the tensor power method.
We prove rates of convergence for our estimator
and provide an empirical evaluation illustrating
its strengths relative to local optimization (EM).

%  Spectral algorithms for latent variable models have seen considerable
%  recent interest for being efficient consistent estimators of model
%  parameters. These algorithms make few if any assumptions about the
%  generative process of the data, while providing a polynomial sample
%  and computational complexity. We present a new spectral algorithm for
%  a discriminative model, a mixture of linear regressions and show that
%  it can recover the regression coefficients with similar polynomial
%  guarantees. We evaluate the algorithm on linearly and non-linearly
%  generated data, as well as on a motion tracking task and compare it's
%  characteristics with an E-M algorithm.
\end{abstract}

### 2013-02-14 doc/icml2013/model.tex
updates

The mixture of linear regressions model describes response variables $y
\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
Each mode is independently described by regression coefficients
$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
the generative process for the response variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

Our learning framework can be described as follows; we observe $N$
samples, $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$ and would like to
recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
\pi_K)$. Note that we do not observe $h$. If we did, the optimal
solution for $\beta_k$ would be the regressor found using least squares
on the set of points, $\{(x_i, y_i) | h_i = k \}$.

>>>>

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
The generation of $y$ given $x$ involves two steps:
(i) draw the mixture component $h \in [K]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_K)$;
and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
variance $\sigma^2$.
%Given $x$, the generative process for the response
%variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi), \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
\end{eqnarray*}
%where $h$ is a latent variable corresponding to the mode.
The parameters of the model are $\theta = (\pi, B, \sigma^2)$,
where $\pi \in \Re^d$ are the mixture proportions,
$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$,
are the regression coefficients,
and $\sigma^2$ is the variance of the observation noise.
The learning problem is stated as follows:
given $N$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{N}, \yn{N})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

%and would like to
%recover the regressors, $B = [\beta_1 \mid \beta_2 \mid \cdots \mid
%\beta_K]$, and the mixture probabilities, $\pi = (\pi_1, \pi_2, \dots,
%\pi_K)$.

%Note that we do not observe $h$. If we did, the optimal
%solution for $\beta_k$ would be the regressor found using least squares
%on the set of points, $\{(x_i, y_i) | h_i = k \}$.

% Typically people use EM, but the maximum marginal likelihood is
% non-convex.
The estimation problem is difficult primarily because the mixture
components $h$ are unobserved, rendering the marginal likelihood
a non-convex function of $\theta$.
Typically, the parameters are learned using
expectation maximization (EM) \cite{VieleTong2002},
which suffers from local optima
(we show this empirically in \sectionref{sec:evaluation}).
%We provide several instances where this leads
%to poor local optima in \sectionref{sec:evaluation}.

This model was used for XXX.

TODO: move to later
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The mixture of linear regressions model is an instance of
mixture of experts \cite{jacobs91},
%wherein the components can be SVMs, GPs, regression trees and other estimators.
%In practice, it is
in which the mixture proportions are allowed to depend on $x$.
We have not considered these extensions in our treatment of the model.
